---
title: "viz_i"
author: "Jianing Chen"
date: "2024-09-26"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") %>%
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create my first scatterplot ever.

```{r}
ggplot(weather_df,aes(x= tmin, y = tmax)) +
  geom_point()
```

New approach, same plot.

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y= tmax)) +
  geom_point()
```
Save and edit a plot project.

```{r}
weather_plot = 
 weather_df %>%
   ggplot(aes(x=tmin, y=tmax))
 
weather_plot + geom_point()
```
## Advanced scatterplot

Start with the same one and make it fancy !

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point() + 
  geom_smooth(se= FALSE)
```

What about the `aes` placement ...?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color= name)) +
  geom_smooth()
```

Let's facet some things !!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y=tmax, color = name)) +
  geom_point(alpha = .2) +
  geom_smooth(se= FALSE)+
  facet_grid(.~name)
  ##facet_grid(name~.)

```

Let's combine some elements and try a new plot.

```{r}
weather_df %>%
  ggplot(aes(x = date, y=tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5)+
  geom_smooth(se = FALSE)+
    facet_grid(.~name)
```

## Some small notes

How many geoms have to exist?

You can have whaterver geoms you want

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_smooth(se = FALSE)
```
 
 You can use a neat geom !
 
```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax))+
  ##geom_hex()
  ##geom_bin2d()
  geom_density2d()+
  geom_point(alpha = .3)
```
## Univariate plots

Histograms are really great.

```{r}
weather_df %>%
  ggplot(aes(x=tmin))+
  geom_histogram()
```


Can we add color ..

```{r}
weather_df %>%
  ##ggplot(aes(x=tmin, color = name))+
  ##geom_histogram()
ggplot(aes(x=tmin, fill = name))+
  geom_histogram(position = "dodge")
##+ facet_grid(.~name)
```

Let's try a new geometry !

```{r}
weather_df %>%
  ggplot(aes(x=tmin, fill= name))+
  geom_density(alpha = .3)
```

What about boxplots ?

```{r}
weather_df %>%
  ##ggplot(aes(y=tmin))+
  ##geom_boxplot()
  ggplot(aes(x=name, y=tmin))+
  geom_boxplot()
```

Trendy plots :)

```{r}
weather_df%>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
  stat_summary()
##geom_violin()
```

Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and Embed

Let's save a scatterplot.

```{r}
weather_plot = 
  weather_df %>%
  ggplot(aes(x=tmin, y= tmax, color = name))+ 
  geom_point(alpha = .5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5 )
```

What about embedding...

```{r}
weather_plot
```

Embed at different size.

```{r, fig,width=12, fig.height = 6}
weather_plot
```

